Rheologic constraints on the upper mantle from five years of postseismic deformation following the El Mayor-Cucapah earthquake

### 2016-06-30 JGRRevision/ElMayor.tex
made some revisions

\item A Zener or Burgers upper mantle can explain far-field deformation while near-field transience can be described with afterslip
\end{keypoints}

@@

Five years of postseismic deformation following the  Mw=7.2 El Mayor-Cucapah earthquake reveals transient deformation that decays back to its pre-earthquake trend after about three years at epicentral distances greater than ${\sim}200$ km. At shorter distances, the rapid transience decays to a sustained rate which exceeds its pre-earthquake trend.  We attempt to determine the mechanisms driving this deformation, where we consider afterslip at seismogenic depths and viscoelastic relaxation in the lower crust and upper mantle as candidate mechanisms.  We find that early, rapid, near-field deformation can be explained with afterslip on the fault that ruptured coseismically, while the later, sustained, near-field deformation can be explained with either continued afterslip in an effectively elastic lower crust, or lower crustal viscoelastic relaxation with a steady-state viscosity of ${\sim}10^{19}$ Pa s. The trend in far-field deformation is best explained with a transient viscosity of ${\sim}10^{18}$ Pa s in the upper mantle. We argue that a transient rheology in the mantle is preferable over a Maxwell rheology because it better predicts the decay in postseismic deformation, and also because it does not conflict with the generally higher, steady-state viscosities inferred from studies of geophysical processes occurring over longer time-scales.
\end{abstract}

@@

Previous studies which have modeled postseismic deformation following the El Mayor-Cucapah earthquake include \citet{Pollitz2012}, \citet{Gonzalez-ortega2014}, \citet{Spinler2015}, and \citet{Rollins2015}. Of these studies, \citet{Gonzalez-ortega2014} and \citet{Rollins2015} have attempted to describe the postseismic deformation with afterslip in an elastic half-space.  \citet{Gonzalez-ortega2014} described five months of postseismic deformation, observed by InSAR and GPS within ${\sim}50$ km of the rupture, with afterslip and contraction on the coseismically ruptured fault. \citet{Gonzalez-ortega2014} noted that their preferred model underestimated the GPS displacements for stations $\gtrsim 25$ km from the rupture and suggested that it could be the result of unmodeled viscoelastic relaxation.  Using only continuous GPS stations, which are mostly north of the rupture zone, \citet{Rollins2015} found that three years of postseismic deformation can be adequately explained by afterslip, albeit with an implausibly large amount of slip inferred on the least constrained, southern-most fault segment. Here, we suggest the afterslip inferred by \citet{Rollins2015} may have been acting as a proxy for distributed relaxation in the upper mantle.

@@

In this study, we perform a kinematic inversion for fault slip, allowing it to persist throughout the postseismic period, while simultaneously estimating the viscosity of the lower crust and upper mantle.  We create an initial model of the fault slip and effective viscosity necessary to describe early postseismic deformation using the method described in \citet{Hines2016}.  This method utilizes a first-order approximation of surface deformation resulting from viscoelastic relaxation which is only applicable to the early postseismic period. In this case, our initial model describes the first 0.8 years of postseismic deformation following the El Mayor-Cucapah earthquake.  We then use the inferred effective viscosity structure from the initial model to create a suite of postseismic models to explain the five years of postseismic data available to date.  Of the suite of models tested, we find that postseismic deformation following the El Mayor-Cucapah earthquake can be explained with a combination of afterslip on a fault segment running through the Sierra Cucapah and a Zener rheology in the upper mantle with a transient viscosity that decays from $6\times10^{18}$ Pa s to $1\times10^{18}$ Pa s at 120 km depth.

@@

\section{Data Processing}\label{sec:Data}
We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) for stations within a 400 km radius about the El Mayor-Cucapah epicenter. We collectively describe the coseismic and postseismic displacements resulting from the El Mayor-Cucapah earthquake as $u_\mathrm{post}(t)$.  We consider the GPS position time series $u_\mathrm{obs}(t)$ to be the combination of $u_\mathrm{post}(t)$, secular tectonic deformation, annual and semi-annual oscillations, and coseismic offsets from significant earthquakes over the time span of this study.  The June 14, 2010, Mw=5.8 Ocotillo earthquake and the Brawley swarm, which included an Mw=5.5 and an Mw=5.4 event on August 26, 2012, (Figure \ref{fig:ContextMap}) are the only earthquakes that produced noticeable displacements in any of the time series.  We treat the displacements resulting from the Brawley swarm as a single event because the daily solutions provided by UNAVCO cannot resolve the separate events.  Although the Ocotillo earthquake had its own series of aftershocks \citep{Hauksson2011}, neither the Ocotillo earthquake nor the Brawley swarm produced detectable postseismic deformation.  We model displacements resulting from these events with only a Heaviside function, $H(t)$, describing the coseismic offsets.  We then model $u_\mathrm{obs}(t)$ as

@@

\begin{equation}\label{TimeSeriesModel}
  \begin{split}  
    u_\mathrm{pred}(t) = &u_\mathrm{post}(t)H(t-t_\mathrm{emc}) + c_0 + c_1t + \\
                         &c_2\sin(2\pi t) + c_3\cos(2\pi t) + c_4\sin(4\pi t) + c_5\cos(4\pi t) + \\
                         &c_6H(t-t_\mathrm{oc}) + c_7H(t-t_\mathrm{bs}).
  \end{split}
\end{equation}
In the above equations, $t_\mathrm{emc}$, $t_\mathrm{oc}$ and $t_\mathrm{bs}$ are the times of the El Mayor-Cucapah earthquake, Ocotillo earthquake, and the Brawley swarm, respectively, $c_0$ through $c_7$ are unknown coefficients, and $\epsilon$ is the observation noise. We only estimate jumps associated with the Ocotillo earthquake and Brawley swarm for stations within 40 km of their epicenters.

Stations which recorded displacements that clearly cannot be described by the aforementioned processes are not included in our analysis. This includes stations in the Los Angeles basin, where anthropogenic deformation can be larger than the postseismic signal that we are trying to estimate \citep{Bawden2001,Argus2005}. In order to ensure an accurate estimation of the secular deformation, we only use stations that were installed at least six months prior to El Mayor-Cucapah earthquake. Several GPS stations were installed after the El Mayor-Cucapah earthquake to improve the spatial resolution of postseismic deformation \citep{Spinler2015}.  Although it would be possible to subtract secular velocities derived from elastic block models \citep[e.g.][]{Meade2005} from velocities recorded at the newly installed stations to get an estimate of postseismic velocities, we do not do so here.  We use coseismic and postseismic displacements, rather than velocities, in our inverse method described in Section \ref{sec:Model}. We use displacements because estimating velocities from an already noisy displacement time series can introduce significant uncertainties depending on exactly how the estimation is done.  This choice prevents us from using the newly installed stations in Baja California for our analysis.

The October 16, 1999, Mw=7.1 Hector Mine earthquake, which occurred ${\sim}270$ km north of the El Mayor-Cucapah epicenter, produced transient postseismic deformation which we do not wish to model, either mechanically or through empirical line fitting.  We thus restrict our analysis to deformation observed six years after the Hector Mine earthquake, which is when postseismic velocities at sites proximal to the Hector Mine epicenter are approximately constant \citep{Savage2009}. When appraising our model fit in Section \ref{sec:Model}, we see some systematic residuals in the vicinity of the Hector Mine epicenter, which may be the result of errors in the assumption that the trend in Hector Mine postseismic deformation is linear after six years.

@@

\begin{equation}
    \mathbf{X}(t) = [u_\mathrm{post}(t),\dot u_\mathrm{post}(t), c_0, ..., c_7]
\end{equation}
and eq. (\ref{TimeSeriesModel}) is the observation function which maps the state vector to the GPS observations. We initiate the Kalman filter by assuming a prior estimate of $\mathbf{X}(t)$ at the first time epoch, denoted $\mathbf{X}_{1|0}$, which has a sufficiently large covariance, denoted $\mathbf{\Sigma}_{1|0}$, to effectively make our prior uninformed.  For each time epoch $t_i$, Bayesian linear regression is used to incorporate GPS derived estimates of displacement with our prior estimate of the state $\mathbf{X}_{i|i-1}$ to form a posterior estimate of the state $\mathbf{X}_{i|i}$, which has covariance $\mathbf{\Sigma}_{i|i}$.  We then use the posterior estimate of the state at time $t_i$ to form a prior estimate of the state at time $t_{i+1}$ through the transition function

@@

The smoothness of $u_\mathrm{post}(t)$ is controlled by the chosen value of $\sigma^2$, which describes how rapidly we expect postseismic displacements to vary over time.  Setting $\sigma^2$ equal to zero will effectively result in modeling $u_\mathrm{post}(t)$ as a straight line which is insufficient to describe the expected transient behavior in postseismic deformation. The other end member, where $\sigma^2$ is infinitely large, will result in $u_\mathrm{pred}(t)$ over fitting the data. While one can use a maximum likelihood based approach for picking $\sigma^2$ \citep[e.g.][]{Segall1997}, we instead take a subjective approach and choose a value for $\sigma^2$ that is just large enough to faithfully describe the observed deformation at the most near-field station in our study, P496, which exhibits the most rapid changes in velocity. This ensures that $\sigma^2$ will be sufficiently large so that our estimate of $u_\mathrm{post}(t)$ does not smooth out potentially valuable postseismic signal at the remaining stations. We find that using $\sigma^2 = 0.05 \mathrm{m}^2 / \mathrm{yr}^3$ adequately describe all but the first week of postseismic deformation at station P496, which gets incorporated into our estimate of coseismic displacements (Figure \ref{fig:P496}).  We assume that the first week of deformation is overwhelmingly the result of afterslip and this afterslip is lumped into our estimates of coseismic slip. \citet{Freed2007a} noted that postseismic deformation can be observed at distances greater than 200 km from the Hector Mine earthquake and, after filtering the time series for stations up to 400 km from the El Mayor-Cucapah epicenter, we also clearly see far reaching postseismic transient deformation (Figure \ref{fig:P619}).

@@

We show in Figures \ref{fig:NearField} and \ref{fig:FarField} the near and far-field postseismic displacements accumulated over the time intervals 0-1 years, 1-3 years, and 3-5 years, as well as the coseismic displacements.  Stations at epicentral distances beyond ${\sim}200$ km have an elevated rate of deformation for the first three years following the earthquake.  This far-field deformation has a southward trend along the direction of the El Mayor-Cucapah P axis and a similar eastward trend can be seen in the few far-field stations in Arizona, located along the T axis.  After three years, the trend in far-field postseismic deformation is barely perceptible.  Most far-field stations display an initial subsidence for the first year after the El Mayor-Cucapah earthquake followed by continued uplift.  This trend in vertical deformation can be observed in all three of the quadrants where postseismic data is available, which means that the vertical deformation does not exhibit an anti-symmetric quadrant pattern, as would be expected for postseismic processes.  Although we use vertical deformation in our analysis in Section \ref{sec:Model},  we do not put an emphasis on trying to describe the vertical deformation as it likely does not have postseismic origins.

\begin{figure}
\includegraphics[scale=0.9]{Figures/MapViewNearField}
\centering 
\caption{Near-field coseismic and cumulative postseismic displacements over the indicated time periods (black) as well as predicted displacements for our preferred model from Section \ref{sec:FullInversion} (green).  Observed vertical deformation is shown as an interpolated field and predicted vertical displacements are shown within the circles at the base of each vector.  Note that the interpolant is not well constrained in Mexico where there is no data available.}
\label{fig:NearField}
\end{figure}

@@

\begin{equation}\label{eq:ElasticForward}
  u_\mathrm{pred}(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi, 
\end{equation}           
where $F$ denotes the fault and $g(x,\xi)$ is the elastic Green's function describing displacement at surface position $x$ resulting from slip at $\xi$ on the fault.  We estimate coseismic slip and the rate of afterslip over the time intervals 0.0-0.125, 0.125-0.25, 0.5-1.0, 1.0-2.0, 2.0-3.0, 3.0-4.0, and 4.0-5.0 years after the earthquake.  Each fault segment is discretized into roughly 4 km by 4 km patches and we estimate a strike-slip and dip component of slip for each patch. We impose that the direction of slip and slip rate are within 45$^\circ$ of right-lateral. We also add zeroth-order Tikhonov regularization so that our solution for $s$ satisfies

@@

Our coseismic slip and afterslip solutions are shown in Figure \ref{fig:ElasticSlip}.  As with \citet{Rollins2015}, we find that a large amount of afterslip on the southern fault segment is required to explain the observations. The potency of our inferred coseismic slip is $3.2\times10^9\ \mathrm{m}^3$, equivalent to a Mw=7.28 earthquake when assuming a shear modulus of 32 GPa.  The potency of our inferred cumulative five years of afterslip is $6.1\times10^9\ \mathrm{m}^3$, equivalent to a Mw=7.46 earthquake, which is unrealistically large if we consider afterslip to be driven by coseismically induced stresses.  Figure \ref{fig:RecordSectionMain} shows the time series for the observed and predicted postseismic displacements at stations along the El Mayor-Cucapah P axis.  We show the radial component of displacements with respect to the El Mayor-Cucapah epicenter and we also rescale the displacements so that the difference between the minimum and maximum observed displacements are the same for each station.  Our elastic slip model accurately describes near-field postseismic deformation and systematically underestimates postseismic deformation at epicentral distances ${\gtrsim}150$ km.  When the fault segments used in the inversion are extended down to 30 km depth, rather than 15 km, the systematic far-field residuals are smaller but remain apparent. Because an elastic model requires an unrealistic amount of afterslip and is unable to predict far-field deformation, we move on to consider viscoelastic models in the next section.

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/RecordSectionFinal}
\caption{Scaled radial component of postseismic displacements. Downward motion indicates that the station is moving toward the El Mayor-Cucapah epicenter.  Displacement time series are scaled so that the minimum and maximum observed values lie on the grid lines.  The observed postseismic displacements, $u_\mathrm{obs}$ are shown in black with gray indicating the 68\% confidence interval.  The displacements predicted by the best fitting elastic model $u_\mathrm{pred}$ are shown in red.  The blue and green lines are the predicted postseismic displacements for the models discussed in Section 3.3. The blue lines show the predicted displacements for the model with a Maxwell viscoelastic lower crust and upper mantle.  The green line shows the predicted displacements for the model with a Maxwell viscoelastic lower crust and a Zener viscoelastic upper mantle.  The effective viscosities are the same for both models and are shown in Figure \ref{fig:EffectiveViscosity}.}
\label{fig:RecordSectionMain}
\end{figure}

@@

In order to greatly simplify the inverse problem, we use the method described in \citet{Hines2016} to constrain an initial effective viscosity structure from the early postseismic deformation.  Our method utilizes the fact that coseismic stresses throughout the crust and upper mantle depend on the instantaneous elastic parameters and are independent of the viscoelastic parameters which we wish to estimate. Immediately following an earthquake, each parcel will have a strain rate that is proportional to the coseismic stress and inversely proportional to the parcel's effective viscosity, 
$\eta_\mathrm{eff}$.  Using one-dimensional viscoelastic models, we define the effective viscosity as

@@

\begin{equation}\label{ObjectiveFunction}
 \mathrm{min}_{s,\eta_\mathrm{eff}} \left(\left|\left|
 \frac{u_\mathrm{pred}(s,\eta_\mathrm{eff}) - u_\mathrm{post}}
 {\mathbf{\sigma_\mathrm{post}}}\right|\right|_2^2 + 
 \lambda_s||s||_2^2 + 
 \lambda_\eta||\nabla \eta_{\mathrm{eff}}^{-1}||_2^2\right),
\end{equation} 
where $u_\mathrm{post}$ consists of the first 0.8 years of postseismic deformation and $u_\mathrm{pred}$ are the predicted displacements from eq. (\ref{ApproxForward}).  Due to inherent non-uniqueness, we have added zeroth-order Tikhonov regularization to estimates of $s$ and second-order Tikhonov regularization to estimates of effective fluidity $\eta_\mathrm{eff}^{-1}$. The degree to which we impose the regularization on slip and fluidity is controlled by the penalty parameters $\lambda_s$ and $\lambda_\eta$, which are chosen with trade-off curves.  Our goal here is to get a prior constraint on $\eta_{\mathrm{eff}}$ to minimize the amount of searching we have to do when describing the postseismic deformation over the full five years, which we do in Section \ref{sec:FullInversion}.  Estimates of $s$ made here will not be used in Section \ref{sec:FullInversion}, and so the motivation behind adding regularization to $s$ is to ensure that the slip driving viscoelastic relaxation in eq. (\ref{ApproxForward}) is sensible.

Our estimated effective viscosities, and corresponding fluidities, are shown in Figure \ref{fig:EffectiveViscosity}.  Although fluidity is rarely used in geophysical literature, eq. (\ref{InitialRate}) is linear with respect to fluidity and so the fluidity indicates the amplitude of the viscoelastic signal coming from each layer.  We also show the 95\% confidence intervals for estimated viscosities and fluidities but we note that the uncertainties tend to be underestimated due to the added regularization \citep{Aster2011}.  A robust feature that we see is that the largest jump in fluidity is at 60 km depth, which is consistent with the range of lithosphere-asthenosphere boundary depths inferred by \citet{Lekic2011}. This transitional depth is also consistent with the the viscosity structure required to explain far-field postseismic deformation following the Hector Mine earthquake \citep{Freed2007a}. We find that the viscosity below 60 km depth needs to be ${\sim}10^{18}$ Pa s to describe the early rate of postseismic deformation at far-field stations while the lower crust and uppermost mantle need to be relatively stronger.  The viscosity of the lower crust is the least well constrained as there is no evidence of relaxation in that layer, meaning that it is effectively elastic over the first 0.8 years after the earthquake.

>>>>

\item Far-field postseismic deformation can be explained with a Zener or Burgers rheology upper mantle

\end{keypoints}

@@

We analyze five years of Southern California GPS data following the Mw=7.2 El Mayor-Cucapah earthquake.  We observed transient postseismic deformation which persists for three years at epicentral distances greater than ${\sim}200$ km.  In the near-field, rapid postseismic transience decays to a sustained rate which exceeds its pre-earthquake trend.  We attempt to determine the mechanisms driving this deformation, where we consider afterslip at seismogenic depths and viscoelastic relaxation in the lower crust and upper mantle as candidate mechanisms.  We find that early, rapid, near-field deformation can be explained with afterslip on the fault that ruptured coseismically, while the later, sustained, near-field deformation can be explained with either continued afterslip in an effectively elastic lower crust, or lower crustal viscoelastic relaxation with a steady-state viscosity of ${\sim}10^{19}$ Pa s. The trend in far-field deformation is best explained with a transient viscosity of ${\sim}10^{18}$ Pa s in the upper mantle. We argue that a transient rheology in the mantle is preferable over a Maxwell rheology because it better predicts the decay in postseismic deformation, and also because it does not conflict with the generally higher, steady-state viscosities inferred from studies of geophysical processes occurring over longer time-scales.
\end{abstract}

@@

Previous studies which have modeled postseismic deformation following the El Mayor-Cucapah earthquake include \citet{Pollitz2012}, \citet{Gonzalez-ortega2014}, \citet{Spinler2015}, and \citet{Rollins2015}. Of these studies, \citet{Gonzalez-ortega2014} and \citet{Rollins2015} have attempted to describe the postseismic deformation with afterslip in an elastic half-space.  \citet{Gonzalez-ortega2014} described five months of postseismic deformation, observed by InSAR and GPS stations within ${\sim}50$ km of the rupture, with afterslip and contraction on the coseismically ruptured fault. \citet{Gonzalez-ortega2014} noted that their preferred model underestimated the GPS displacements for stations $\gtrsim 25$ km from the rupture and suggested that it could be the result of unmodeled viscoelastic relaxation.  Using only continuous GPS stations, which are mostly north of the rupture zone, \citet{Rollins2015} found that three years of postseismic deformation can be adequately explained by afterslip, albeit with an implausibly large amount of slip inferred on the least constrained, southern-most fault segment. Here, we suggest the afterslip inferred by \citet{Rollins2015} may have been acting as a proxy for distributed relaxation in the upper mantle.

@@

In this study, we perform a kinematic inversion for fault slip, allowing it to persist throughout the postseismic period, while simultaneously estimating the viscosity of the lower crust and upper mantle.  We create an initial model of the fault slip and effective viscosity necessary to describe early postseismic deformation using the method described in \citet{Hines2016}.  This method uses a first-order approximation of surface deformation resulting from viscoelastic relaxation which is only applicable to the early postseismic period. In this case, our initial model describes the first 0.8 years of postseismic deformation following the El Mayor-Cucapah earthquake.  We then use the inferred effective viscosity structure from the initial model to create a suite of postseismic models to explain the five years of postseismic data available to date.  Of the suite of models tested, we find that postseismic deformation following the El Mayor-Cucapah earthquake can be explained with a combination of afterslip on a fault segment running through the Sierra Cucapah and a Zener rheology in the upper mantle with a transient viscosity on the order of $10^{18}$ Pa s.

@@

\section{Data Processing}\label{sec:Data}
We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) for stations within a 400 km radius about the El Mayor-Cucapah epicenter. We collectively describe the coseismic and postseismic displacements resulting from the El Mayor-Cucapah earthquake as $u_\mathrm{post}(t)$.  We consider the GPS position time series, $u_\mathrm{obs}(t)$, to be the combination of $u_\mathrm{post}(t)$, secular tectonic deformation, annual and semi-annual oscillations, and coseismic offsets from significant earthquakes over the time span of this study.  The June 14, 2010, Mw=5.8 Ocotillo earthquake and the Brawley swarm, which included an Mw=5.5 and an Mw=5.4 event on August 26, 2012, (Figure \ref{fig:ContextMap}) are the only earthquakes that produced noticeable displacements in any of the time series.  We treat the displacements resulting from the Brawley swarm as a single event because the daily solutions provided by UNAVCO cannot resolve the separate events.  Although the Ocotillo earthquake had its own series of aftershocks \citep{Hauksson2011}, neither the Ocotillo earthquake nor the Brawley swarm produced detectable postseismic deformation.  We model displacements resulting from these events with only a Heaviside function, $H(t)$, describing the coseismic offsets.  We then model $u_\mathrm{obs}(t)$ as

@@

\begin{equation}\label{TimeSeriesModel}
  \begin{split}  
    u_\mathrm{pred}(t) = &u_\mathrm{post}(t)H(t-t_\mathrm{emc}) + c_0 + c_1t + \\
                         &c_2\sin(2\pi t) + c_3\cos(2\pi t) + c_4\sin(4\pi t) + c_5\cos(4\pi t) + \\
                         &c_6H(t-t_\mathrm{oc}) + c_7H(t-t_\mathrm{bs}).
  \end{split}
\end{equation}
In the above equations, $t_\mathrm{emc}$, $t_\mathrm{oc}$ and $t_\mathrm{bs}$ are the times of the El Mayor-Cucapah earthquake, Ocotillo earthquake, and the Brawley swarm, respectively, $c_0$ through $c_7$ are unknown coefficients, and $\epsilon$ is the observation noise.  We are using years as our unit of time which makes $c_2$ through $c_5$ the coefficients for annual and semi-annual oscillations.  We only estimate jumps associated with the Ocotillo earthquake and Brawley swarm for stations within 40 km of their epicenters.

Stations which recorded displacements that clearly cannot be described by the aforementioned processes are not included in our analysis. This includes stations in the Los Angeles basin, where anthropogenic deformation can be larger than the postseismic signal that we are trying to estimate \citep{Bawden2001,Argus2005}. In order to ensure an accurate estimation of the secular deformation, we only use stations that were installed at least six months prior to El Mayor-Cucapah earthquake even though several GPS stations were installed after the earthquake to get better coverage of the postseismic deformation field \citep{Spinler2015}.  It would be possible to subtract secular velocities derived from elastic block models \citep[e.g.][]{Meade2005} from velocities recorded at the newly installed stations to get an estimate of postseismic velocities at those stations. However, estimating velocities from an already noisy displacement time series can introduce significant uncertainties depending on exactly how the estimation is done.  We therefore use coseismic and postseismic displacements, rather than velocities, in our inverse method described in Section \ref{sec:Model}. This choice prevents us from using the newly installed stations for our analysis.

The October 16, 1999, Mw=7.1 Hector Mine earthquake, which occurred ${\sim}270$ km north of the El Mayor-Cucapah epicenter, produced transient postseismic deformation which we do not wish to model, either mechanically or through empirical line fitting.  We thus restrict our analysis to deformation observed six years after the Hector Mine earthquake, which is when postseismic velocities at sites near the Hector Mine epicenter are approximately constant \citep{Savage2009}. When appraising our model fit in Section \ref{sec:Model}, we see some systematic residuals in the vicinity of the Hector Mine epicenter, which may be the result of errors in the assumption that the trend in Hector Mine postseismic deformation is linear after six years.

@@

\begin{equation}
    \mathbf{X}(t) = [u_\mathrm{post}(t),\dot u_\mathrm{post}(t), c_0, ..., c_7]
\end{equation}
and eq. (\ref{TimeSeriesModel}) is the observation function which maps the state vector to the GPS observations. We initiate the Kalman filter by assuming a prior estimate of $\mathbf{X}(t)$ at the first time epoch, denoted $\mathbf{X}_{1|0}$, which has a sufficiently large covariance, denoted $\mathbf{\Sigma}_{1|0}$, to effectively make our prior uninformed.  For each time epoch, $t_i$, Bayesian linear regression is used to incorporate GPS derived estimates of displacement with our prior estimate of the state, $\mathbf{X}_{i|i-1}$, to form a posterior estimate of the state, $\mathbf{X}_{i|i}$, which has covariance $\mathbf{\Sigma}_{i|i}$.  We then use the posterior estimate of the state at time $t_i$ to form a prior estimate of the state at time $t_{i+1}$ through the transition function

@@

The smoothness of $u_\mathrm{post}(t)$ is controlled by the chosen value of $\sigma^2$, which describes how rapidly we expect postseismic displacements to vary over time.  Setting $\sigma^2$ equal to zero will effectively result in modeling $u_\mathrm{post}(t)$ as a straight line which is insufficient to describe the expected transient behavior in postseismic deformation. The other end member, where $\sigma^2$ is infinitely large, will result in $u_\mathrm{pred}(t)$ over fitting the data. While one can use a maximum likelihood based approach for picking $\sigma^2$ \citep[e.g.][]{Segall1997}, we instead take a subjective approach and choose a value for $\sigma^2$ that is just large enough to faithfully describe the observed deformation at the most near-field station in our study, P496, which exhibits the most rapid changes in velocity. This ensures that $\sigma^2$ will be sufficiently large so that our estimate of $u_\mathrm{post}(t)$ does not smooth out potentially valuable postseismic signal at the remaining stations. We find that using $\sigma^2 = 0.05 \mathrm{m}^2 / \mathrm{yr}^3$ adequately describe all but the first week of postseismic deformation at station P496, which gets incorporated into our estimate of coseismic displacements (Figure \ref{fig:P496}).  We assume that the first week of deformation is overwhelmingly the result of afterslip and this afterslip is lumped into our estimates of coseismic slip.  We include an example of estimating $u_\mathrm{post}(t)$ for a far-field station, P619, which is about 359 km from the El Mayor-Cucapah epicenter (Figure \ref{fig:P619}).  At station P619, along with all the other stations in the Mojave region, there is a south-trending postseismic transience that persists for the first three years after the El Mayor-Cucapah earthquake.  Postseismic deformation that extends to these epicentral distances has also been observed after the Hector Mine earthquake \citep{Freed2007a}.

@@

We show in Figures \ref{fig:NearField} and \ref{fig:FarField} the near and far-field postseismic displacements accumulated over the time intervals 0-1 years, 1-3 years, and 3-5 years, as well as the coseismic displacements.  Stations at epicentral distances beyond ${\sim}200$ km have an elevated rate of deformation for the first three years following the earthquake.  This far-field deformation is trending southward at a rate of a few millimeters per year along the direction of the El Mayor-Cucapah P axis.  A similar eastward trend can be seen in the few far-field stations in Arizona, located along the T axis.  After three years, the trend in far-field postseismic deformation is barely perceptible.  Most far-field stations display an initial subsidence for the first year after the El Mayor-Cucapah earthquake followed by continued uplift.  This trend in vertical deformation can be observed in all three of the quadrants where postseismic data is available, which means that the vertical deformation does not exhibit an anti-symmetric quadrant pattern, as would be expected for postseismic processes.  Although we use vertical deformation in our analysis in Section \ref{sec:Model},  we do not put an emphasis on trying to describe the vertical deformation as it likely does not have postseismic origins.

\begin{figure}
\includegraphics[scale=0.9]{Figures/MapViewNearField}
\centering 
\caption{Near-field coseismic and cumulative postseismic displacements over the indicated time periods (black) as well as predicted displacements for our preferred model from Section \ref{sec:FullInversion} (green).  Observed vertical deformation is shown as an interpolated field and predicted vertical displacements are shown within the green circles.  Note that the interpolant is not well constrained in Mexico where there is no data available.}
\label{fig:NearField}
\end{figure}

@@

\begin{equation}\label{eq:ElasticForward}
  u_\mathrm{pred}(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi, 
\end{equation}           
where $F$ denotes the fault and $g(x,\xi)$ is the elastic Green's function describing displacement at surface position $x$ resulting from slip at $\xi$ on the fault.  We estimate coseismic slip and the rate of afterslip over the time intervals 0.0-0.125, 0.125-0.25, 0.25-0.5, 0.5-1.0, 1.0-2.0, 2.0-3.0, 3.0-4.0, and 4.0-5.0 years after the earthquake.  Each fault segment is discretized into roughly 4 km by 4 km patches and we estimate a strike-slip and dip component of slip for each patch. We impose that the direction of slip and slip rate are within 45$^\circ$ of right-lateral. We also add zeroth-order Tikhonov regularization so that our solution for $s$ satisfies

@@

Our coseismic slip and afterslip solutions are shown in Figure \ref{fig:ElasticSlip}.  As with \citet{Rollins2015}, we find that a large amount of afterslip on the Indiviso fault segment is required to explain the observations. The potency of our inferred coseismic slip is $3.2\times10^9\ \mathrm{m}^3$, equivalent to a Mw=7.28 earthquake when assuming a shear modulus of 32 GPa.  The potency of our inferred cumulative five years of afterslip is $6.1\times10^9\ \mathrm{m}^3$, equivalent to a Mw=7.46 earthquake, which is unrealistically large if we consider afterslip to be driven by coseismically induced stresses.  Figure \ref{fig:RecordSectionMain} shows the time series for the observed and predicted postseismic displacements at stations along the El Mayor-Cucapah P axis.  We show the radial component of displacements with respect to the El Mayor-Cucapah epicenter and we also rescale the displacements so that the difference between the minimum and maximum observed displacements are the same for each station.  Our elastic slip model accurately describes near-field postseismic deformation and systematically underestimates postseismic deformation at epicentral distances ${\gtrsim}150$ km.  When the fault segments used in the inversion are extended down to 30 km depth, rather than 15 km, the systematic far-field residuals are smaller but remain apparent. Because an elastic model requires an unrealistic amount of afterslip and is unable to predict far-field deformation, we move on to consider viscoelastic models in the next section.

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/RecordSectionFinal}
\caption{Scaled radial component of postseismic displacements. Downward motion indicates that the station is moving toward the El Mayor-Cucapah epicenter.  Displacement time series are scaled so that the minimum and maximum observed values lie on the grid lines.  The observed postseismic displacements, $u_\mathrm{post}$ are shown in black with gray indicating the 68\% confidence interval.  The displacements predicted by the best fitting elastic model $u_\mathrm{pred}$ are shown in red.  The blue and green lines are the predicted postseismic displacements for the models discussed in Section 3.3. The blue lines show the predicted displacements for the model with a Maxwell viscoelastic lower crust and upper mantle.  The green line shows the predicted displacements for the model with a Maxwell viscoelastic lower crust and a Zener viscoelastic upper mantle.  The effective viscosities are the same for both models and are shown in Figure \ref{fig:EffectiveViscosity}.}
\label{fig:RecordSectionMain}
\end{figure}

@@

In order to greatly simplify the inverse problem, we use the method described in \citet{Hines2016} to constrain an initial effective viscosity structure from the early postseismic deformation.  Our method uses the fact that coseismic stresses throughout the crust and upper mantle depend on the instantaneous elastic parameters and are independent of the viscoelastic parameters which we wish to estimate. Immediately following an earthquake, each parcel will have a strain rate that is proportional to the coseismic stress and inversely proportional to the parcel's effective viscosity, $\eta_\mathrm{eff}$.  Using one-dimensional rheologic models, we define the effective viscosity as

@@

\begin{equation}\label{ObjectiveFunction}
 \mathrm{min}_{s,\eta_\mathrm{eff}} \left(\left|\left|
 \frac{u_\mathrm{pred}(s,\eta_\mathrm{eff}) - u_\mathrm{post}}
 {\mathbf{\sigma_\mathrm{post}}}\right|\right|_2^2 + 
 \lambda_s||s||_2^2 + 
 \lambda_\eta||\nabla \eta_{\mathrm{eff}}^{-1}||_2^2\right),
\end{equation} 
where $u_\mathrm{post}$ consists of the first 0.8 years of postseismic deformation and $u_\mathrm{pred}$ are the predicted displacements from eq. (\ref{ApproxForward}).  Due to inherent non-uniqueness, we have added zeroth-order Tikhonov regularization to estimates of $s$ and second-order Tikhonov regularization to estimates of effective fluidity $\eta_\mathrm{eff}^{-1}$. The degree to which we impose the regularization on slip and fluidity is controlled by the penalty parameters $\lambda_s$ and $\lambda_\eta$, which are chosen with trade-off curves (Figure S1).  Our goal here is to get a prior constraint on $\eta_{\mathrm{eff}}$ to minimize the amount of searching we have to do when describing the postseismic deformation over the full five years, which we do in Section \ref{sec:FullInversion}.  Estimates of $s$ made here will not be used in Section \ref{sec:FullInversion}, and so the motivation behind adding regularization to $s$ is to ensure that the slip driving viscoelastic relaxation in eq. (\ref{ApproxForward}) is sensible.

Our initial estimate for coseismic slip and cumulative afterslip over the first 0.8 years after the El Mayor-Cucapah earthquake are shown in Figure \ref{fig:InitialSlip}.  Similar to our elastic slip model from Section \ref{sec:ElasticInversion}, a significant amount of right-lateral and normal coseismic slip is inferred to be on the Sierra Cucapah segment. Our coseismic slip solution on the Sierra Cucapah segment is consistent with field studies \citep{Fletcher2014} and the model from \citet{Wei2011}.  Our inferred slip on the Indiviso fault segment differs from \citet{Wei2011} because the GPS data used in this study is not capable of resolving the spatial distribution of fault slip on that segment (Figure S2).  The potency of inferred coseismic slip is $3.3\times 10^{9}\ \mathrm{m}^3$, which is also about the same as that inferred from Section \ref{sec:ElasticInversion}. The present inference of afterslip on the Indiviso fault is significantly less than what was found in the Section \ref{sec:ElasticInversion} where we did not account for viscoelasticity. When fault slip is simultaneously estimated with viscosity, the potency of inferred afterslip over the first 0.8 years after the earthquake is $0.85\times 10^9\ \mathrm{m}^3$, compared to $3.5\times10^{9}\ \mathrm{m}^3$ when we assume the crust and upper mantle are elastic.  The significant amount of afterslip inferred on the Indiviso fault seems to be compensating for unmodeled viscoelastic relaxation at depths greater than $60$ km.  The fact that there is still an appreciable amount of afterslip inferred on the Indiviso fault raises the question of whether it is compensating for viscoelastic relaxation that is more localized than what we allow for since we only estimate depth dependent variations in viscosity.
